---
title: "p8105_hw1_jy3360"
author: "Candice Yu"
date: "2023-09-14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use the code below to download the a package containing the penguins dataset:

```{r}
#install.packages("moderndive")
library("moderndive")
data("early_january_weather")
```

# Problem 1

Write a short description of the dataset using inline R code; accessing the dataset help file can be informative. In your discussion, please include:

-   the variables in this dataset, including names / values of important variables
-   the size of the dataset (using nrow and ncol)
-   the mean temperature

```{r}
#?early_january_weather
#View(early_january_weather)
#summary(early_january_weather)
length(names(early_january_weather))
names(early_january_weather)
nrow(early_january_weather)
mean(early_january_weather$temp)
```

-   Here are `r length(names(early_january_weather))` variables, which are `r names(early_january_weather)`.
-   The number of row is `r nrow(early_january_weather)` and the number of columns is `r ncol(early_january_weather)`.
-   The mean temperature is `r round(mean(early_january_weather$temp, na.rm=TRUE), digits = 2)`F.
-   Names / values of important variables are in the table below. 

```{r}
skimr::skim(early_january_weather)
```


Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot.

```{r}
library("ggplot2")
ggplot(data = early_january_weather, 
       mapping = aes(x = time_hour, 
                     y = temp,
                     color = humid)) +
  geom_point(na.rm = TRUE)
```

- The plot shows that the temperature has an increasing trend from Jan 1 to Jan 15. 
- The percentage of humidity increased from Jan 10 to Jan 15. 
- The range of temperature is similar for each day. 


Export your scatterplot to your project directory using ggsave.
```{r}
ggsave("scatterplot.pdf")
```

# Problem 2

Create a data frame comprised of:
- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r}
library(tidyverse)
df_2 <- tibble(
    var_norm = rnorm(n = 10),
    var_logical = var_norm > 0,
    var_character = c("Amy", "John", "Mike", "Jojo", "Rachel", "Luke", "Michael", "Kim", "Emma", "Jenny"),
    var_factor = factor(c("low", "high", "moderate", "high", "moderate", "moderate", "high", "low", "high", "low"))
  )
df_2
```

Question: Try to take the mean of each variable in your dataframe. What works and what doesn’t?

Hint: for now, to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.
```{r}
mean(pull(df_2, var_norm))
mean(pull(df_2, var_logical))
mean(pull(df_2, var_character))
mean(pull(df_2, var_factor))
```

- It works when calculating the mean of numeric and logical vectors, but it doesn’t work when calculating the mean of character and factor vectors. 


Question: In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

```{r conversion, eval = FALSE}
as.numeric(pull(df_2, var_logical))
as.numeric(pull(df_2, var_character))
as.numeric(pull(df_2, var_factor))
```

- The results show that as.numeric function does not work for character vectors, but works for logical and factor vectors. 
- R cannot identify the character data into numeric data and so the result is "NA". 
- When I convert the logical variable into numeric, it converts "TRUE" to 1 and "FALSE" to 0. 
- Although the factor variable has not been assigned with order, R gives them order alphabetically, and assign them with value from 1 to 3. 
- We can take the mean for logical and factor vectors but not for character vectors. 


